An Adolescent Mini-Neptune in the Kepler Field

### 2021-06-16 paper/ms.tex
update ms

\begin{abstract}
  Kepler\,1627A is an active G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
  data from the Gaia satellite with rotation periods from TESS, we
  show that Kepler\,1627 is a member of the $35\pm 10$\,Myr old
  $\delta$~Lyr cluster.  Our analysis advances Kepler\,1627Ab as the
  youngest planet with a precise age detected by the main Kepler
  mission.  New ages from the combination of Gaia and TESS offer the
  opportunity to significantly expand the census of age-dated planets
  -- we offer a literature concatenation of young stars to support
  this effort.  Future studies of \pn\ may also help clarify how the
  orbits and atmospheres of the mini-Neptune planets evolve.
\end{abstract}

@@

@@

Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
transit may hint at some small degree of unfitted signal -- in the
sense that the observations are systematically high in the first half
of transit, and low in the second half.  
One possible explanation could be that we are seeing the effects of
gravity darkening (e.g., CITE Masuda 2015).
The expected amplitude of the effect is XXX, based on CITE.
However, there is also a dip of comparable amplitude shortly before
the transit. 
Given the uncertainties in the stellar variability model, we expect
that these patterns may correspond to more noise than signal.

>>>>

\begin{abstract}
  Kepler\,1627A is an active G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we
  show that Kepler\,1627 is a member of the $35\pm XX$\,Myr old
  $\delta$~Lyr cluster.  Our analysis advances Kepler\,1627Ab as the
  youngest planet with a precise age found by the main Kepler
  mission.  
  Leveraging the corresponding photometry, we explore the possibility
  that star-planet interactions induce flares in the system, and find
  evidence that starspot crossings distort the average transit shape.
  More broadly though, the merging of Gaia and TESS data offers an
  opportunity to significantly expand the census of age-dated planets;
  we provide a literature concatenation of young stars to support this
  effort.
  In the future, spectroscopic studies of \pn\ may also help clarify
  how the orbits and atmospheres of the mini-Neptune planets evolve.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f9.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Transit model residuals through time}.  
    {\it Left:}
		Phase-folded transit of Kepler 1627b, with stellar
    variability removed, binned by Kepler quarter.
    Black points are binned to 20 minute intervals.
    The 1-$\sigma$ model uncertainties and the maximum {\it a
    posteriori} model are shown as the faint purple band, and the dark
    purple line.
    {\it Right:}
    As on the left, with the transit removed.
    Quarters 6 and 7 show a consistent deviation in the second half of
    the transit.
		\label{fig:phasequarter}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Transit model residuals through time (Part 2)}.  
    {\it Left:}
		Phase-folded transit of Kepler 1627b, with stellar
    variability removed, binned by year of observation.
    Points and models are as in Figure~\ref{fig:phasequarter}.
    {\it Right:}
    As on the left, with the transit removed.
		\label{fig:phaseyear}
	}
\end{figure*}

@@

Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
transit may hint at some small degree of unfitted signal -- in the
sense that the observations are systematically high in the first half
of transit, and low in the second half.  This is the phase of maximal
scatter during each orbital period (TODO: in vs out of transit
scatter).  The amplitude of the anomaly is $\approx$30\,ppm.

To explore the origin of the anomaly, we binned the Kepler data over
quarters (Figure~\ref{fig:phasequarter}) and years
(Figure~\ref{fig:phaseyear}).  In Figure~\ref{fig:phasequarter}
Quarter 6 shows the strongest asymmetry out of any of the quarters: a
deviation of about 3\,ppt from expectation.  Quarter 7 shows an
anomaly at roughly the same transit phase.  Year 2 correspondingly
shows the strongest anomaly out of any year in
Figure~\ref{fig:phaseyear}.

We considered three possible explanations for the anomaly: gravity
darkening, transit timing variations, and spot-crossing events.

Gravity darkening (e.g., CITE Masuda 2015) is based on the premise
that the rapidly rotating star becomes oblate, and brighter near the
poles than the equator.  The fractional shape change due to gravity
darkening is on the order of $(P_{\rm break}/P_{\rm rot})^2$, for
$P_{\rm break}$ the break-up rotation period, and $P_{\rm rot}$ the
rotation period.  Using the parameters from Table~\ref{tab:posterior},
this yields an expected 1.6\% distortion of the $\approx$1.8\,ppt
transit depth: {\it i.e.}, an absolute deviation of $\approx$3\,ppm.
This is smaller than the observed anomaly by an order of magnitude,
and therefore seems unlikely.

The scenario of transit timing variations (TTVs) producing the
asymmetry seems unlikely, since the analysis by
\citet{holczer_transit_2016} implies that any such variations in \sn\
need to be less than of order minutes.  Figures~\ref{fig:phasequarter}
and~\ref{fig:phaseyear} also provide little evidence in support of
this possibility.

The final possibility is that of starspot crossings.  Given the high
expected spot-covering fractions for young stars ({\it e.g.},
\citealt{morris_relationship_2020}, \citealt{plavchan_planet_2020}),
\pn\ may cross spot groups on the stellar surface in projection.
Spot-crossing anomalies often reach amplitudes exceeding 100\,ppm
({\it e.g.}, CITE Dai 2018, maybe Southworth's recent stuff).  For our
system,  $P_{\rm orb}/P_{\rm rot} \approx 2.76$.  This means that
every 4 transits (and 11 stellar rotations), the planet crosses over
roughly the same stellar longitude.  Over a given quarter, this could
occur at most three times, assuming the spot groups are persistent.

This could provide a plausible path to creating the distorted transit
signal, based on the amplitudes involved.  However, the typical S/N
per Kepler transit is $\approx8$, and so the individual spot-crossing
events cannot be resolved.  {\it A priori}, one would also expect the
spot-crossing events to not have a preferred phase, so that they would
average out over the $\approx200$ Kepler transits.  Nonetheless, given
the available data, it is our best explanation for what we believe is
real additional scatter during the transits, relative to the
out-of-transit observations.
